---
title: "zoonotic_capacity"
output: html_document
---

```{r setup try2, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#knitr::knit("zoonotic_capacity.Rmd", output = "html_document")
```

#1. 
#####install packages needed  
```{r packages, include = FALSE}

library(tidyr)
print(Sys.time())
pkgTest <- function(x)
{
  if (x %in% rownames(installed.packages()) == FALSE) {
    install.packages(x, dependencies= TRUE, repos='http://cran.us.r-project.org')    
  }
  library(x, character.only = TRUE)
}
neededPackages <- c( "ggplot2",  "plyr", "foreach","broom", "remotes", "stringr",  "data.table",  "Hmisc", "caret", "ggforce",  "gbm", "caret", "Matrix", "pdp", "caTools", "ROCR", "dplyr", "dismo", "doSNOW",  "tidyverse",  "patchwork", "gtable", "magrittr",  "fulltext",  "operator.tools", "stringi", "tidyr", "knitr", "parallel")
for (package in neededPackages){pkgTest(package)}

```

#2. 
#folders -- create output and figure folders and delete files in them if they already exist
```{r folder}
mainDir = c("output", "figure")
for (a in 1:length(mainDir)){
  if (!dir.exists(file.path(mainDir[a]))){
    dir.create(file.path(mainDir[a]))
  } else { 
    list_files = list.files(mainDir[a])
    file.remove(paste0(mainDir[a], "/", list_files))
  }  
}

performance_out = NULL#this will store performance metrics
save(performance_out, file = "output/performance_out.Rdata")

pred_sum_out = NULL
save(pred_sum_out, file = "output/pred_sum_out.Rdata")

```

#3.
##cores
```{r, include = FALSE}
check_cores = detectCores(TRUE)#find how many cores are available for parallel processing
cores = check_cores/4#note: this may need to be changed to use no more than half the number of cores available
cl <- makeCluster(cores, "SOCK", timeout = 60)  
registerDoSNOW(cl)

#if getting error related to dopar, nodes
# stopCluster(cl)
# rm(cl)
  
```

#4. 
#Define data and model. #This is the zoonotic capacity model that had the highest accuracy, including only wild species in the training and test data. 
```{r wild_only -129_imputed zoonotic capacity}
source("code/remove_temp_data.R")#remove any temporary data 
source("code/wild_only_zoonotic_capacity_include_forstrat_conservative_deviance_2.R")#this script defines the model and removes fields with near-zero variation 

```

#5.
#This code chunk carries out grid search. 
```{r grid_search wild_only zoonotic capacity }

source("code/R_run_gridSearch.R")#run gridSearch function to evaluate alternative hyperparameter combinations and how they perform 
source("code/deviance_plots_all.R")#Make a subplot for the deviance curve for each hyperparameter combination
#Look at the deviance curve, together with the test accuracy statistic (AUC), for each hyperparameter combination. Pick the best parameters, which have highest test AUC while also having a relatively flat test deviance curve and having optimal trees fewer than the maximum trees. 

#Based on this evaluation, these are the hyperparameters we chose to use in modeling: 
eta = 0.0001
max_depth = 2
n.minobsinnode = 5

source("code/deviance_plot_defined.R")#As a double-check that we picked the hyperparameter combo we intended to, plot deviance curve for just these hyperparameters. 
```

#6. Run bootstrapping of boosted regression tree modeling with chosen set of hyperparameters. 
```{r bootstrapping wild_only zoonotic capacity}
##run scripts
source("code/scripts.R")
```

#7.
#make predictions once, using all available data. do one shot of making predictions with all of the available data and with the same set of hyperparameters as used in "conservative_depth_2"
```{r predictions_once}
source("code/wild_only_zoonotic_capacity_include_forstrat_cons_dev_2_one_iteration.R")#define model
source("code/run_bootstrapGBM_predictions_one_iteration.R")
```


##################################################################################################################################
#8. 
###Below here are code chunks for running every other model formulation that we evaluated and found to be relatively uninformative. The code is commented out. The code can be run by taking out the comments. For more info on these alternative models see Supplementary Methods and Supplementary Table 4 (available here: https://doi.org/10.25390/caryinstitute.c.5293339)

##do analysis, mammals, imputed, wild and non-wild, haddock_infected_and_below (-129)
```{r wild_non_wild_zoonotic_capacity}
#source("code/wild_non_wild_zoonotic_capacity.R")

```

##mammals, imputed, haddock_score_mean
```{r mammals_haddock.R}
#source("code/mammals_haddock.R")

```

##mammals, imputed, AA_30_negative
```{r mammals, imputed, AA_30_negative}
#source("code/mammals_AA_haddock.R")

```

#all #vertebrates, zoonotic capacity
```{r vertebrates_zoonotic_capacity.R}
#source("code/vertebrates_zoonotic_capacity.R")

```

##all vertebrates, haddock_score_mean
```{r vertebrates_haddock.R}
#source("code/vertebrates_haddock.R")
```

##vertebrates, AA 30
```{r vertebrates_AA_haddock.R}
#source("code/vertebrates_AA_haddock.R")
```

##write outputs -- performance_out and DF_merge_out. Make csv file that has performance metrics for all models run.  
```{r performance}
#source("code/remove_temp_data.R")
#load("output/performance_out.Rdata")
#load("output/pred_sum_out.Rdata")
#write.csv(performance_out, paste0("output/performance_models", ".csv"))
#save(pred_sum_out, file = "output/pred_sum_out.Rdata")
```

##find out how long it takes to run
```{r}
print(Sys.time())
```

